The Atala PRISM Protocol

### 2023-01-04 prism-backend/research/prism-protocol/did_atala.tex
Updated steps in the proof.

We now prove that \RealPKIDIDAtala $\xi$-UC-emulates \IdealGPKIDID, in the
presence of global subroutines \IdealGclock, \IdealGRO, and \IdealGdledger.
Given that (multiple) global subroutines are involved, the proving process
requires several steps, which we summarise first, and then specify in detail. In
this process, we rely on recent results in \cite{bch+20,bhz21} as well as,
naturally, on the UC framework \cite{canetti00}. In a nutshell, the steps we
take are as follows (here, we are intentionally imprecise on occasions, in
favour of giving a high-level idea of the overall approach):

\todo{Some diagram would probably help much here.}
\begin{enumerate}
\item \emph{(Merging global subroutines into \gAtala)} Following the strategy
  in \cite{bhz21}, we combine all the global subroutines needed in our protocols
  into a single (equivalent) one. While merely ``bureaucratic'', this
  facilitates the analysis in subsequent steps. We refer to this combined
  global shared subroutine as \gAtala.
\item \emph{(\gAtala is subroutine respecting)} We show that only main ITIs
  in \gAtala accept input from ITIs not in \gAtala, no ITI within \gAtala passes
  input to ITIs not in \gAtala, no ITI in \gAtala expects subroutine output from
  other ITIs outside of \gAtala, and all subroutine outputs produced within
  \gAtala are sent to the outside world via \gAtala's main ITIs.
\item \emph{(\RealPKIDIDAtala is \gAtala-subroutine respecting)} We show that
  \RealPKIDIDAtala is also subroutine respecting, as just described for \gAtala,
  except actually for calls to \gAtala. This means that ITIs within
  \RealPKIDIDAtala may communicate directly with machines in \gAtala without
  needing to ``pass through'' the main ITIs of \RealPKIDIDAtala, and can
  communicate with them even if they are not part of \RealPKIDIDAtala's session.
\item \emph{(\IdealGPKIDID is \gAtala-subroutine respecting)} Same as before,
  but with \IdealGPKIDID.
\item \emph{(\gAtala is \RealPKIDIDAtala-regular)} We show that ITIs in \gAtala
  do not invoke new ITIs of \RealPKIDIDAtala as subroutines, nor call or expect
  subroutine output from machines of \RealPKIDIDAtala.
\item \emph{(\gAtala is \IdealGPKIDID-regular)} Same as before, but with
  \IdealGPKIDID.
\item \emph{(Executions of $\UCM[\RealPKIDIDAtala,\gAtala]$ are indistinguishable
    from execution of \RealPKIDIDAtala in the presence of \gAtala)} Given the
  previous preconditions, \cite[Proposition 3.4]{bch+20} ensures that we can
  safely apply a transformation $\UCM[\cdot,\cdot]$ to \RealPKIDIDAtala, which
  allows to leverage the techniques from the plain UC framework.
\item \emph{(Executions of $\UCM[\IdealGPKIDID,\gAtala]$ are indistinguishable
    from execution of \IdealGPKIDID in the presence of \gAtala)} Same as before,
  but with \IdealGPKIDID.
\item \emph{(\RealPKIDIDAtala $\xi$-UC-emulates \IdealGPKIDID in the presence
    of \gAtala)} By \cite[Definition 3.1]{bch+20}, if $\UCM[\RealPKIDIDAtala,
  \gAtala]$ $\xi$-UC-emulates $\UCM[\IdealGPKIDID,\gAtala]$, then
  \RealPKIDIDAtala $\xi$-UC-emulates \IdealGPKIDID, in the presence
  of \gAtala.
\end{enumerate}

@@

\paragraph{1. Merging global subroutines into \gAtala.} This transformation
is straightforward, and can be done as in \cite[Definition 4.1]{bhz21}. Namely,
let $\gamma_1 \gets \IdealGledger$, $\gamma_2 \gets \IdealGRO$, and $\gamma_3
\gets \IdealGclock$. Then, we define \gAtala to be a functionality that simply
accepts inputs of type $(\uccmd{query},\sid,i,x)$, and makes a query to
$\gamma_i$, with input $x$ on session \sid -- e.g., a query $(\uccmd{query},
\sid,2,(\uccmd{HasQuery},m))$ results in a query $(\uccmd{HashQuery},\sid,m)$ to
\IdealGRO.
%
This allows us to directly apply the results in \cite{bch+20}. \todo{Later, we
  will get back to arguing that the usual candidates for UC-realising the
  individual $\gamma_i$'s meet certain properties that make them suitable for
  UC-realising the ``merged'' functionality \gAtala. Thus, our main results
  apply when considering real instantiations of the global subroutines, rather
  than their idealised counterparts.}

\paragraph{2. \gAtala is subroutine respecting.}

\commentwho{Jesus}{Not removing what follows for now,  because it will be
  useful. But keep on the proof process stated in the previous enumeration.}

>>>>

We now prove that \RealPKIDIDAtala $\xi$-UC-emulates \IdealGPKIDID, in the
presence of global subroutines \IdealGclock, \IdealGRO, and \IdealGdledger,
as well as local digital signatures functionality \IdealFSig.
Given that (multiple) global subroutines are involved, the proving process
requires several steps, which we summarise first, and then specify in detail. In
this process, we rely on recent results in \cite{bch+20,bhz21} as well as,
naturally, on the UC framework \cite{canetti00}. In a nutshell, the steps we
take are as follows (here, we are intentionally imprecise on occasions, in
favour of giving a high-level idea of the overall approach):

\todo{Some diagram would probably help much here.}
\begin{enumerate}
\item \emph{(Merging global subroutines into \gAtala)} Following the strategy
  in \cite{bhz21}, we combine all the global subroutines needed in our protocols
  into a single (equivalent) one. While merely ``bureaucratic'', this
  facilitates the analysis in subsequent steps. We refer to this combined
  global shared subroutine as \gAtala.
\item \emph{(\gAtala is subroutine respecting)} We show that only main ITIs
  in \gAtala accept input from ITIs not in \gAtala, no ITI within \gAtala passes
  input to ITIs not in \gAtala, no ITI in \gAtala expects subroutine output from
  other ITIs outside of \gAtala, and all subroutine outputs produced within
  \gAtala are sent to the outside world via \gAtala's main ITIs.
\item \emph{(\RealPKIDIDAtala is \gAtala-subroutine respecting)} We show that
  \RealPKIDIDAtala is also subroutine respecting, as just described for \gAtala,
  except actually for calls to \gAtala. This means that ITIs within
  \RealPKIDIDAtala may communicate directly with machines in \gAtala without
  needing to ``pass through'' the main ITIs of \RealPKIDIDAtala, and can
  communicate with them even if they are not part of \RealPKIDIDAtala's session.
\item \emph{(\IdealGPKIDID is \gAtala-subroutine respecting)} Same as before,
  but with \IdealGPKIDID.
\item \emph{(\gAtala is \RealPKIDIDAtala-regular)} We show that ITIs in \gAtala
  do not invoke new ITIs of \RealPKIDIDAtala as subroutines, nor call or expect
  subroutine output from machines of \RealPKIDIDAtala.
\item \emph{(\gAtala is \IdealGPKIDID-regular)} Same as before, but with
  \IdealGPKIDID.
\item \emph{(Executions of $\UCM[\RealPKIDIDAtala,\gAtala]$ are indistinguishable
    from executions of \RealPKIDIDAtala in the presence of \gAtala)} Given the
  previous preconditions, \cite[Proposition 3.4]{bch+20} ensures that we can
  safely apply a transformation $\UCM[\cdot,\cdot]$ to \RealPKIDIDAtala, which
  allows to leverage the techniques from the plain UC framework.
\item \emph{(Executions of $\UCM[\IdealGPKIDID,\gAtala]$ are indistinguishable
    from executions of \IdealGPKIDID in the presence of \gAtala)} Same as before,
  but with \IdealGPKIDID.
\item \emph{(\RealPKIDIDAtala $\xi$-UC-emulates \IdealGPKIDID in the presence
    of \gAtala)} By \cite[Definition 3.1]{bch+20}, if $\UCM[\RealPKIDIDAtala,
  \gAtala]$ $\xi$-UC-emulates $\UCM[\IdealGPKIDID,\gAtala]$, then
  \RealPKIDIDAtala $\xi$-UC-emulates \IdealGPKIDID, in the presence
  of \gAtala.
\end{enumerate}

@@

\paragraph{1. Merging global subroutines into \gAtala.} This transformation
is straightforward, and can be done as in \cite[Definition 4.1]{bhz21}. Namely,
let $\gamma_1 \gets \IdealGledger$, $\gamma_2 \gets \IdealGRO$, and $\gamma_3
\gets \IdealGclock$. Then, we define \gAtala to be a functionality that simply
accepts inputs of type $(\uccmd{query},\sid,i,x)$, and makes a query to
$\gamma_i$, with input $x$ on session \sid -- e.g., a query $(\uccmd{query},
\sid,2,(\uccmd{HashQuery},m))$ results in a query $(\uccmd{HashQuery},\sid,m)$
to \IdealGRO.
%
This allows us to directly apply the results in \cite{bch+20}. \todo{Later, we
  will get back to arguing that the usual candidates for UC-realising the
  individual $\gamma_i$'s meet certain properties that make them suitable for
  UC-realising the ``merged'' functionality \gAtala. Thus, our main results
  apply when considering real instantiations of the global subroutines, rather
  than their idealised counterparts.}

\paragraph{2. \gAtala is subroutine respecting.} We start by observing that
all of the $\gamma_i$'s that compose \gAtala are subroutine respecting.
Concretely, on the one hand:

\begin{itemize}
\item \IdealGclock (resp. \IdealGRO) is the only sub-party in the protocol
  session \sid, and it only accepts input from the main (dummy) parties in
  $\IdealP_{\IdealGclock}$ (resp. $\IdealP_{\IdealGRO}$) in \sid.
\item The main (dummy) parties of session \sid in $\IdealP_{\IdealGclock}$
  (resp. $\IdealP_{\IdealGRO}$) ignore subroutine outputs from parties other
  than \IdealGclock (resp. \IdealGRO)'s on session \sid. Similarly,
  \IdealGclock (resp. \IdealGRO) ignores subroutine outputs.
\item \IdealGclock (resp. \IdealGRO) in \sid only passes subroutine output to
  the main (dummy) parties in \sid.
\item Main (dummy) parties in session \sid of $\IdealP_{\IdealGclock}$ (resp.
  $\IdealP_{\IdealGRO}$), and \IdealGclock (resp. \IdealGRO)  don't pass input
  to parties not in \sid. 
\end{itemize}

On the other hand, \IdealGdledger ($\gamma_1$) is also subroutine respecting
except for calls to \IdealGclock ($\gamma_3$). Concretely, for each session
of the ideal protocol $\IdealP_{\IdealGdledger}$ of \IdealGdledger:

\begin{itemize}
\item Excluding calls to \IdealGclock, \IdealGdledger is the only sub-party in
  the protocol session \sid, and it only accepts input from the main (dummy)
  parties in $\IdealP_{\IdealGdledger}$ in \sid.
\item The main (dummy) parties of session \sid in $\IdealP_{\IdealGdledger}$
  ignore subroutine outputs from parties other than \IdealGdledger's on
  session \sid. Also, \IdealGdledger ignores subroutine outputs from any
  party, with the exception from parties in \IdealGclock.
\item \IdealGdledger in \sid only passes subroutine output to the main (dummy)
  parties of $\IdealP_{\IdealGdledger}$, and to \IdealGclock, in \sid.
\item Main (dummy) parties in session \sid of $\IdealP_{\IdealGdledger}$ and
  \IdealGdledger don't pass input to parties not in \sid, with the exception
  of parties in $\IdealP_{\IdealGclock}$.
\end{itemize}

In \gAtala, the main (dummy) parties of each $\gamma_i$ become main (dummy)
parties of \gAtala. Thus, from the previous, it is easy to see that \gAtala is
subroutine respecting. \todo{Perhaps add a figure here -- like the one in my
  notes.}

\paragraph{3. \RealPKIDIDAtala is \gAtala-subroutine respecting.}
The only parties with which \RealPKIDIDAtala communicates (concretely, via
passing subroutine requests, or receiving subroutine responses) that do not
share the same \sid as \RealPKIDIDAtala, are \IdealGclock, \IdealGRO, and
\IdealGdledger -- all of them bundled within \gAtala. Thus, \RealPKIDIDAtala is
\gAtala-subroutine respecting.

\paragraph{4. \IdealGPKIDID is \gAtala-subroutine respecting.}
As in the previous step, the only parties with which \IdealGPKIDID communicates
that do not share the same \sid as \IdealGPKIDID, is \IdealGclock, which is
bundled within \gAtala. Thus, \IdealGPKIDID is also \gAtala-subroutine
respecting.

\paragraph{5. \gAtala is \RealPKIDIDAtala-regular.} No main party of
\IdealGdledger, \IdealGRO, or \IdealGclock (respectively, $\gamma_1$,
$\gamma_2$, and $\gamma_3$), invokes a new ITI of \RealPKIDIDAtala via the
subroutine output tape or has an ITI with code \RealPKIDIDAtala as subsidiary.
Since \gAtala bundles all $\gamma_i$'s, the same applies to \gAtala, and
thus \gAtala is \RealPKIDIDAtala-regular.

\paragraph{6. \gAtala is \IdealGPKIDID-regular.} No main party of
\IdealGdledger, \IdealGRO, or \IdealGclock (respectively, $\gamma_1$,
$\gamma_2$, and $\gamma_3$), invokes a new ITI of \IdealGPKIDID via the
subroutine output tape or has an ITI with code \IdealGPKIDID as subsidiary.
Since \gAtala bundles all $\gamma_i$'s, the same applies to \gAtala, and
thus \gAtala is \IdealGPKIDID-regular.

\paragraph{7. Executions of $\UCM[\RealPKIDIDAtala,\gAtala]$ are
  indistinguishable from executions of \RealPKIDIDAtala in the presence of
  \gAtala.}

\paragraph{8. Executions of $\UCM[\IdealGPKIDID,\gAtala]$ are indistinguishable
  from executions of \IdealGPKIDID in the presence of \gAtala.}

\paragraph{9. \RealPKIDIDAtala $\xi$-UC-emulates \IdealGPKIDID in the presence
  of \gAtala.}

\commentwho{Jesus}{Not removing what follows for now,  because it may/will be
  useful. But keep on the proof process stated in the previous enumeration.}
